---
title: 'HighChart visualizations'
author: "Chris Olande"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: hide
---

# Highcharter R Visualization Examples

This document demonstrates various chart types using the `highcharter` package for creating interactive visualizations in R.

## Setup
```{r setup}
librarian::shelf(highcharter, forecast, quantmod, TSstudio, tidyverse, zoo)
```

## 1. Scatter Plot - Palmer Penguins
Creates an interactive scatter plot showing the relationship between flipper length and bill length, grouped by penguin species.
```{r penguin_scatter}
data(penguins, package = "palmerpenguins")
hchart(
  penguins,
  "scatter",
  hcaes(x = flipper_length_mm, y = bill_length_mm, group = species)
)
```

## 2. Histogram - Penguin Bill Length Distribution
Displays the distribution of penguin bill lengths as a histogram.
```{r penguin_histogram}
hchart(penguins$bill_length_mm, name = "Bill Length", color = "#17b8b6")
```

## 3. Time Series Forecast - Air Passengers
Shows ARIMA forecast for the classic AirPassengers dataset with confidence intervals.
```{r air_passengers_forecast}
airforecast <- forecast(auto.arima(AirPassengers), level = 95)
hchart(airforecast)
```

## 4. Stock Chart - Google & Amazon
Creates a financial stock chart with candlestick (OHLC) data for Google and Amazon stocks.
```{r stock_chart_multiple}
x <- getSymbols("GOOG", auto.assign = FALSE)
y <- getSymbols("AMZN", auto.assign = FALSE)

highchart(type = "stock") |>
  hc_add_series(x) |>
  hc_add_series(y, type = "ohlc")
```

## 5. World Map - GNI Per Capita
Interactive world map showing Gross National Income per capita with logarithmic color scale.
```{r world_map_gni}
data(GNI2014, package = "treemap")
hcmap(
  "custom/world-robinson-lowres",
  data = GNI2014,
  name = "Gross national income per capita",
  value = "GNI",
  borderWidth = 0,
  nullColor = "#d3d3d3",
  joinBy = c("iso-a3", "iso3")
) |>
  hc_colorAxis(
    stops = color_stops(colors = viridisLite::inferno(10, begin = 0.1)),
    type = "logarithmic"
  )
```

## 6. Treemap - Car Manufacturers
Treemap visualization showing car manufacturers by count and model diversity.
```{r treemap_cars}
mpgman <- mpg |>
  group_by(manufacturer) |>
  summarise(n = n(), unique = length(unique(model))) |>
  arrange(-n, -unique)

hchart(mpgman, "treemap", hcaes(x = manufacturer, value = n, color = unique))
```

## 7. Multi-line Chart - Economic Indicators
Time series chart showing multiple economic indicators (population, unemployment) over time.
```{r economic_indicators}
data(diamonds, economics_long, package = "ggplot2")
economics_long2 <- economics_long |>
  filter(variable %in% c("pop", "uempmed", "unemploy"))

hchart(economics_long2, "line", hcaes(x = date, y = value01, group = variable))
```

## 8. Histogram - Diamond Prices
Simple histogram of diamond prices from the diamonds dataset.
```{r diamond_price_histogram}
x <- diamonds$price
hchart(x)
```

## 9. Density Plot - Diamond Prices
Area chart showing the density distribution of diamond prices.
```{r diamond_price_density}
hchart(density(x), type = "area", color = "#B71C1C", name = "Price")
```

## 10. Bar Chart - Diamond Cut Quality
Column chart showing the frequency of different diamond cut qualities.
```{r diamond_cut_bar}
x <- diamonds$cut
hchart(x, type = "column")
```

## 11. Time Series - Lake Huron Levels
Line chart of Lake Huron water levels from 1875-1972.
```{r lake_huron_ts}
hchart(LakeHuron, name = "Level") |>
  hc_title(text = "Level of Lake Huron 1875–1972")
```

## 12. STL Decomposition - Air Passengers
Seasonal and trend decomposition of the AirPassengers time series.
```{r stl_decomposition}
x <- stl(log(AirPassengers), "per")
hchart(x)
```

## 13. Alternative Time Series Decomposition
Using TSstudio package for time series decomposition visualization.
```{r tsstudio_decomposition}
ts_decompose(AirPassengers)
```

## 14. ETS Forecast - US Accidental Deaths
Exponential smoothing forecast for US accidental deaths data.
```{r ets_forecast}
x <- forecast(ets(USAccDeaths), h = 48, level = 95)
hchart(x)
```

## 15. Single Stock Chart - Google
Simple stock price chart for Google (GOOG).
```{r single_stock_chart}
x <- getSymbols("GOOG", auto.assign = FALSE)
hchart(x)
```

## 16. Multiple Time Series - Deaths Data
Combined chart showing male and female deaths over time.
```{r multiple_deaths_ts}
x <- cbind(mdeaths, fdeaths)
hchart(x)
```

## 17. PACF Plot - Partial Autocorrelation
Partial autocorrelation function plot for differenced AirPassengers data.
```{r pacf_plot}
x <- pacf(diff(AirPassengers), plot = FALSE)
hchart(x)
```

## 18. PCA Analysis - US Arrests
Principal component analysis visualization of US Arrests data.
```{r pca_analysis}
hchart(princomp(USArrests, cor = TRUE))
```

## 19. Correlation Matrix - Motor Trends
Heatmap showing correlations between variables in the mtcars dataset.
```{r correlation_matrix}
hchart(cor(mtcars))
```

## Additional Chart Examples

### 20. Bubble Chart - Cars Data
```{r bubble_chart_cars}
hchart(
  mtcars,
  "bubble",
  hcaes(x = wt, y = mpg, size = hp, group = cyl)
) |>
  hc_title(text = "Car Weight vs MPG vs Horsepower")
```

### 21. Pie Chart - Transmission Types
```{r pie_chart_transmission}
mtcars_summary <- mtcars |>
  count(am) |>
  mutate(transmission = ifelse(am == 0, "Automatic", "Manual"))

hchart(mtcars_summary, "pie", hcaes(name = transmission, y = n)) |>
  hc_title(text = "Transmission Types in mtcars Dataset")
```

### 22. Area Chart - Stacked Economic Data
```{r area_chart_economic}
economics_wide <- economics |>
  select(date, unemploy, pop) |>
  mutate(
    unemploy_thousands = unemploy / 1000,
    pop_millions = pop / 1000000
  )

highchart() |>
  hc_add_series(
    economics_wide,
    "area",
    hcaes(x = date, y = unemploy_thousands),
    name = "Unemployment (thousands)"
  ) |>
  hc_add_series(
    economics_wide,
    "area",
    hcaes(x = date, y = pop_millions),
    name = "Population (millions)",
    yAxis = 1
  ) |>
  hc_yAxis_multiples(
    list(title = list(text = "Unemployment")),
    list(title = list(text = "Population"), opposite = TRUE)
  ) |>
  hc_title(text = "US Population and Unemployment Over Time")
```

### 23. Box Plot - Diamond Prices by Cut
```{r box_plot_diamonds}
hcboxplot(
  x = diamonds$price,
  var = diamonds$cut,
  name = "Price",
  color = "#2980b9"
) |>
  hc_title(text = "Diamond Prices by Cut Quality")
```

### 24. Polar Chart - Wind Rose
```{r polar_chart_wind}
set.seed(123)
wind_data <- data.frame(
  direction = seq(0, 350, 10),
  speed = abs(rnorm(36, 10, 3))
)

hchart(wind_data, "column", hcaes(x = direction, y = speed), polar = TRUE) |>
  hc_title(text = "Wind Speed by Direction") |>
  hc_pane(startAngle = 0, endAngle = 360)
```

### 25. Gauge Chart - Performance Metric
```{r gauge_chart_performance}
highchart() |>
  hc_chart(type = "solidgauge") |>
  hc_pane(
    startAngle = -90,
    endAngle = 90,
    background = list(
      outerRadius = "100%",
      innerRadius = "60%",
      shape = "arc"
    )
  ) |>
  hc_add_series(
    data = list(85),
    name = "Performance",
    dataLabels = list(
      enabled = TRUE,
      format = "<span style='font-size: 25px'>{y}%</span>"
    )
  ) |>
  hc_yAxis(
    min = 0,
    max = 100,
    title = list(text = "Performance Score")
  ) |>
  hc_title(text = "Performance Dashboard")
```

### 26. Waterfall Chart - Financial Analysis
```{r waterfall_chart_financial}
financial_data <- data.frame(
  category = c("Starting", "Revenue", "Costs", "Taxes", "Ending"),
  value = c(100, 50, -30, -10, 110),
  type = c("start", "positive", "negative", "negative", "end")
)

hchart(financial_data, "waterfall", hcaes(name = category, y = value)) |>
  hc_title(text = "Financial Waterfall Analysis")
```

### 27. Funnel Chart - Conversion Rates
```{r funnel_chart_conversion}
funnel_data <- data.frame(
  stage = c("Visitors", "Leads", "Prospects", "Customers"),
  count = c(1000, 400, 200, 50)
)

hchart(funnel_data, "funnel", hcaes(name = stage, y = count)) |>
  hc_title(text = "Sales Funnel Conversion")
```

## Extended Time Series Examples

### 28. Seasonal Heatmap - Temperature Data
```{r seasonal_heatmap_temperature}
# Create synthetic temperature data
set.seed(42)
temp_data <- expand.grid(
  month = 1:12,
  year = 2015:2024
) |>
  mutate(
    date = as.Date(paste(year, month, "01", sep = "-")),
    temperature = 15 +
      10 * sin(2 * pi * (month - 3) / 12) +
      rnorm(n(), 0, 2),
    month_name = month.name[month]
  )

hchart(
  temp_data,
  "heatmap",
  hcaes(x = year, y = month_name, value = temperature)
) |>
  hc_colorAxis(
    stops = color_stops(colors = c("#3060cf", "#fffbbc", "#c4463a")),
    min = 0,
    max = 30
  ) |>
  hc_title(text = "Monthly Temperature Heatmap (2015-2024)") |>
  hc_xAxis(title = list(text = "Year")) |>
  hc_yAxis(title = list(text = "Month"), reversed = TRUE)
```

### 29. Smoothed Time Series with Confidence Bands
```{r smoothed_ts_confidence}
# Create time series with trend and noise
set.seed(123)
dates <- seq(as.Date("2020-01-01"), as.Date("2024-12-31"), by = "week")
trend <- 100 + 0.5 * seq_along(dates) + 20 * sin(2 * pi * seq_along(dates) / 52)
noise <- rnorm(length(dates), 0, 5)
values <- trend + noise

# Calculate moving average and confidence bands

ma <- rollmean(values, k = 4, fill = NA, align = "center")
upper_band <- ma +
  1.96 * rollapply(values, width = 4, FUN = sd, fill = NA, align = "center")
lower_band <- ma -
  1.96 * rollapply(values, width = 4, FUN = sd, fill = NA, align = "center")

ts_data <- data.frame(
  date = dates,
  observed = values,
  smoothed = ma,
  upper = upper_band,
  lower = lower_band
)

highchart() |>
  hc_add_series(
    ts_data,
    "line",
    hcaes(x = date, y = observed),
    name = "Observed",
    color = "#cccccc",
    lineWidth = 1
  ) |>
  hc_add_series(
    ts_data,
    "arearange",
    hcaes(x = date, low = lower, high = upper),
    name = "Confidence Band",
    color = "#3498db",
    fillOpacity = 0.3
  ) |>
  hc_add_series(
    ts_data,
    "line",
    hcaes(x = date, y = smoothed),
    name = "Smoothed Trend",
    color = "#e74c3c",
    lineWidth = 2
  ) |>
  hc_title(text = "Time Series with Smoothed Trend and Confidence Bands") |>
  hc_xAxis(title = list(text = "Date")) |>
  hc_yAxis(title = list(text = "Value"))
```

### 30. Multiple Stock Comparison with Volume
```{r multiple_stock_comparison}
# Get multiple stocks (if quantmod is available)
stocks <- c("AAPL", "MSFT", "GOOGL")
stock_data <- lapply(stocks, function(symbol) {
  tryCatch(
    {
      getSymbols(symbol, auto.assign = FALSE, from = "2023-01-01")
    },
    error = function(e) NULL
  )
})

if (
  !is.null(stock_data[[1]]) &&
    !is.null(stock_data[[2]]) &&
    !is.null(stock_data[[3]])
) {
  # Create comparison chart
  highchart(type = "stock") |>
    hc_add_series(stock_data[[1]][, 6], name = "AAPL") |>
    hc_add_series(stock_data[[2]][, 6], name = "MSFT") |>
    hc_add_series(stock_data[[3]][, 6], name = "GOOGL") |>
    hc_title(text = "Stock Price Comparison") |>
    hc_rangeSelector(enabled = TRUE)
}
```

### 31. Irregular Time Series - Event Data
```{r irregular_ts_events}
# Create irregular time series (events happening at random intervals)
set.seed(456)
event_dates <- sort(sample(
  seq(as.Date("2023-01-01"), as.Date("2024-12-31"), by = "day"),
  50
))
event_values <- cumsum(rnorm(50, 2, 5))

irregular_data <- data.frame(
  date = event_dates,
  value = event_values,
  event_type = sample(c("Type A", "Type B", "Type C"), 50, replace = TRUE)
)

hchart(
  irregular_data,
  "line",
  hcaes(x = date, y = value, group = event_type)
) |>
  hc_plotOptions(
    line = list(
      marker = list(enabled = TRUE, radius = 4),
      lineWidth = 2
    )
  ) |>
  hc_title(text = "Irregular Time Series - Event Data") |>
  hc_xAxis(title = list(text = "Date")) |>
  hc_yAxis(title = list(text = "Cumulative Value"))
```

### 32. Time Series with Annotations and Flags
```{r ts_annotations_flags}
# Create time series with significant events
set.seed(789)
ts_dates <- seq(as.Date("2020-01-01"), as.Date("2024-01-01"), by = "month")
ts_values <- 100 * cumprod(1 + rnorm(length(ts_dates), 0.005, 0.05))

# Define significant events
events <- data.frame(
  date = as.Date(c("2020-03-15", "2021-06-01", "2022-12-01", "2023-09-15")),
  event = c(
    "Market Crash",
    "Recovery Begins",
    "Policy Change",
    "New Regulation"
  ),
  impact = c("negative", "positive", "neutral", "positive")
)

# Create the main series
ts_chart <- highchart() |>
  hc_add_series(
    data = data.frame(date = ts_dates, value = ts_values),
    type = "line",
    hcaes(x = date, y = value),
    name = "Index Value",
    color = "#2c3e50"
  )

# Add event flags
for (i in 1:nrow(events)) {
  flag_color <- switch(events$impact[i],
    "positive" = "#27ae60",
    "negative" = "#e74c3c",
    "neutral" = "#f39c12"
  )

  ts_chart <- ts_chart |>
    hc_add_series(
      data = list(list(
        x = as.numeric(events$date[i]) * 86400000,
        title = substr(events$event[i], 1, 1)
      )),
      type = "flags",
      name = events$event[i],
      color = flag_color,
      fillColor = flag_color,
      style = list(color = "white")
    )
}

ts_chart |>
  hc_title(text = "Time Series with Event Annotations") |>
  hc_xAxis(title = list(text = "Date")) |>
  hc_yAxis(title = list(text = "Index Value"))
```

### 33. Seasonal Subseries Plot
```{r seasonal_subseries}
# Create seasonal data
seasonal_data <- data.frame(
  date = seq(as.Date("2020-01-01"), as.Date("2023-12-31"), by = "month"),
  sales = 1000 + 500 * sin(2 * pi * (1:48) / 12) + rnorm(48, 0, 100)
) |>
  mutate(
    year = year(date),
    month = month(date),
    month_name = month.name[month]
  )

# Create subseries by month
hchart(seasonal_data, "line", hcaes(x = year, y = sales, group = month_name)) |>
  hc_title(text = "Seasonal Subseries Plot - Monthly Sales by Year") |>
  hc_xAxis(title = list(text = "Year")) |>
  hc_yAxis(title = list(text = "Sales")) |>
  hc_plotOptions(
    line = list(
      marker = list(enabled = TRUE)
    )
  )
```

### 34. Time Series Comparison - Before/After Analysis
```{r before_after_analysis}
# Create before/after intervention data
set.seed(101)
dates <- seq(as.Date("2022-01-01"), as.Date("2024-12-31"), by = "week")
intervention_date <- as.Date("2023-06-01")

# Different patterns before and after intervention
before_after <- ifelse(dates < intervention_date, "Before", "After")
base_trend <- ifelse(dates < intervention_date, 50, 65)
seasonal_effect <- 10 * sin(2 * pi * as.numeric(dates - min(dates)) / 365.25)
noise <- rnorm(length(dates), 0, 5)

values <- base_trend + seasonal_effect + noise

comparison_data <- data.frame(
  date = dates,
  value = values,
  period = before_after
)

hchart(comparison_data, "line", hcaes(x = date, y = value, group = period)) |>
  hc_plotOptions(
    line = list(
      lineWidth = 2,
      marker = list(enabled = FALSE)
    )
  ) |>
  hc_xAxis(
    title = list(text = "Date"),
    plotLines = list(
      list(
        color = "#e74c3c",
        dashStyle = "Dash",
        value = as.numeric(intervention_date) * 86400000,
        width = 2,
        label = list(text = "Intervention", style = list(color = "#e74c3c"))
      )
    )
  ) |>
  hc_yAxis(title = list(text = "Metric Value")) |>
  hc_title(text = "Before/After Intervention Analysis")
```

### 35. Multi-frequency Time Series
```{r multi_frequency_ts}
# Combine daily, weekly, and monthly data
daily_dates <- seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "day")
weekly_dates <- seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "week")
monthly_dates <- seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "month")

multi_freq_data <- list(
  daily = data.frame(
    date = daily_dates,
    value = 100 + cumsum(rnorm(length(daily_dates), 0, 1)),
    frequency = "Daily"
  ),
  weekly = data.frame(
    date = weekly_dates,
    value = 100 + cumsum(rnorm(length(weekly_dates), 0, 2)),
    frequency = "Weekly"
  ),
  monthly = data.frame(
    date = monthly_dates,
    value = 100 + cumsum(rnorm(length(monthly_dates), 0, 5)),
    frequency = "Monthly"
  )
) |>
  bind_rows()

hchart(
  multi_freq_data,
  "line",
  hcaes(x = date, y = value, group = frequency)
) |>
  hc_plotOptions(
    line = list(
      lineWidth = 2
    )
  ) |>
  hc_colors(c("#3498db", "#e74c3c", "#2ecc71")) |>
  hc_title(text = "Multi-frequency Time Series Comparison") |>
  hc_xAxis(title = list(text = "Date")) |>
  hc_yAxis(title = list(text = "Value"))
```

### 36. Regime Change Detection
```{r regime_change_detection}
# Simulate regime changes in time series
set.seed(202)
n_points <- 200
dates <- seq(as.Date("2020-01-01"), length.out = n_points, by = "week")

# Create different regimes
regime1 <- rnorm(60, 10, 2) # Low volatility, mean 10
regime2 <- rnorm(80, 15, 5) # High volatility, mean 15
regime3 <- rnorm(60, 8, 1.5) # Low volatility, mean 8

values <- c(regime1, regime2, regime3)
regimes <- c(rep("Regime 1", 60), rep("Regime 2", 80), rep("Regime 3", 60))

regime_data <- data.frame(
  date = dates,
  value = values,
  regime = regimes
)

hchart(regime_data, "line", hcaes(x = date, y = value, group = regime)) |>
  hc_plotOptions(
    line = list(
      lineWidth = 2,
      marker = list(enabled = FALSE)
    )
  ) |>
  hc_title(text = "Time Series with Regime Changes") |>
  hc_xAxis(title = list(text = "Date")) |>
  hc_yAxis(title = list(text = "Value"))
```

### 37. Time Series Correlation Matrix
```{r ts_correlation_matrix}
# Create multiple correlated time series
set.seed(303)
n_series <- 5
n_points <- 100
dates <- seq(as.Date("2023-01-01"), length.out = n_points, by = "day")

# Generate correlated series
base_series <- cumsum(rnorm(n_points, 0, 1))
corr_matrix <- matrix(
  c(
    1,
    0.8,
    0.6,
    0.3,
    -0.2,
    0.8,
    1,
    0.7,
    0.4,
    -0.1,
    0.6,
    0.7,
    1,
    0.5,
    0.1,
    0.3,
    0.4,
    0.5,
    1,
    0.3,
    -0.2,
    -0.1,
    0.1,
    0.3,
    1
  ),
  nrow = 5
)

# Generate correlated data
series_data <- MASS::mvrnorm(n_points, mu = rep(0, 5), Sigma = corr_matrix)
colnames(series_data) <- paste("Series", LETTERS[1:5])

# Convert to long format for plotting

ts_long <- as.data.frame(series_data) |>
  mutate(date = dates) |>
  pivot_longer(cols = -date, names_to = "series", values_to = "value")

hchart(ts_long, "line", hcaes(x = date, y = value, group = series)) |>
  hc_plotOptions(
    line = list(
      lineWidth = 2,
      marker = list(enabled = FALSE)
    )
  ) |>
  hc_title(text = "Multiple Correlated Time Series") |>
  hc_xAxis(title = list(text = "Date")) |>
  hc_yAxis(title = list(text = "Value"))
```
